Effects of bariatric surgery and weight loss on resting state functional connectivity

### 2021-02-01 manuscript/draft_rsfmri.Rmd
problem with subfloat in float - does not knit

---
title: "Effects of bariatric surgery and weight loss on resting state functional connectivity"
author:
- Hannah Sophie Heinrichs* [1]
- Frauke Beyer* [1][2]
- Kristin Prehn
- Jürgen Ordemann
- Agnes Floel
- A. Veronica Witte [1][2][3]
date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3
  Day Clinic for Cognitive Neurology, University Clinic Leipzig
output:
  bookdown::pdf_book:
    fig_caption: yes
    always_allow_html: true
    number_sections: yes
    toc: yes
    extra_dependencies: "subfig"
bibliography: "bibliography.bib"
editor_options:

  chunk_output_type: console
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

```{r "knit setup", include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
options(knitr.table.format = "latex")
options(knitr.kable.NA = '') # hide NA in tables
# set working directory of knit (RMarkdown) to directory of Rproj
#opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
#setwd(knitr::opts_knit$get("root.dir"))

@@

```{r calc DMN two tp, age and sex,results = 'asis', echo = FALSE}
#### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)

@@

```{qa check DMN model age,sex, echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

```{r DMN2tp results age, sex, caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)

@@

```{r calc DMN two tp, age and sex and mFD,results = 'asis', echo = FALSE}
#### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)

@@

```{qa check DMN model age,sex, mFD, echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

```{r DMN2tp results age sex mFD, caption="DMN connectivity compared bl to fu, adjusted for age, sex and mFD", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "logmFD", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

```{r calc DMN three tp, eval = TRUE, echo = FALSE}
#### Aggregated DMN for three timepoints, adjusted for age, sex, mean FD (Model 1b)
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
anvDMN3tp=anova(R1,R0)
```

```{r DMN3tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

```{r calc rew two tp, age, sex, eval = TRUE, echo = FALSE}
### Reward network for two time points, adjusted for age and sex (Model 1a)

@@

```{qa check Rew model age,sex, echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

```{r Rew2tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

```{r calc rew two tp adjusted, eval = TRUE, echo = FALSE}
#### Reward network for two timepoints, adjusted for age, sex, mean FD (Model 1b)

@@

```{qa check Rew model age,sex,mFD, echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

```{r Rew2tpadj results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","logmFD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

```{r calc rew three tp, eval = TRUE, echo = FALSE}
#### Reward network for three timepoints, adjusted for age, sex, mean FD (Model 1b)

R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC)
anvRew3tp=anova(R1,R0)
```
```{r Rew3tp results, caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

@@

#### Aggregated FC analysis
```{r DMN model, mean, change BMI, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]

R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
R02a = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
anvmeanBMI_adjmFD=anova(R1,R02a)
```
Corresponding to the whole brain results, there was a negative association of mean BMI and aggregated DMN connectivity (p = `r round(p_meanBMI,3)`), regardless of whether we adjusted for mFD.
```{r Rew model, mean, change BMI,, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_Rew_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
```

```{r DMN model mean&change of BMI&logmFD, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_DMN_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
anvmeanFD=anova(R1,R03)
anvwithinFD=anova(R1,R03)

@@

#### Reward network, mean and within variability in BMI and logmFD
```{r Rew model mean&change of BMI&logmFD, eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_Rew_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
R04 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
anvwithinFD=anova(R1,R03)
anvmeanFD=anova(R1,R04)

>>>>

---
title: "Effects of bariatric surgery and weight loss on resting state functional connectivity"
author:
- Hannah Sophie Heinrichs* [1]
- Frauke Beyer* [1][2]
- Kristin Prehn
- Jürgen Ordemann
- Agnes Floel
- A. Veronica Witte [1][2][3]
date: 1 Max-Planck-Institute for Human Cognitive and Brain Sciences, Leipzig \newline
  2 CRC 1052 "Obesity Mechanisms", Subproject A1, University of Leipzig \newline 3
  Day Clinic for Cognitive Neurology, University Clinic Leipzig
output:
  bookdown::pdf_book:
    fig_caption: yes
    always_allow_html: true
    number_sections: yes
    toc: yes
    extra_dependencies: "subfig"
bibliography: "bibliography.bib"
editor_options:
  chunk_output_type: console
header-includes:
- \usepackage[left]{lineno}
- \linenumbers
---

@@

```{r "knit setup", include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE) # fig.height=2, fig.width=4)
options(knitr.table.format = "latex")
options(knitr.kable.NA = '') # hide NA in tables
knitr::knit_engines$get()
# set working directory of knit (RMarkdown) to directory of Rproj
#opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
#setwd(knitr::opts_knit$get("root.dir"))

@@

```{r "calc DMN two tp, age and sex" ,results = 'asis', echo = FALSE}
#### Aggregated DMN for two timepoints, adjusted for age and sex (Model 1a)

@@

```{r "qa check DMN model age,sex", echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

```{r "DMN2tp results age, sex", caption="DMN connectivity compared bl to fu, adjusted for age and sex", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)

@@

```{r "calc DMN two tp, age and sex and mFD", results = 'asis', echo = FALSE}
#### Aggregated DMN for two timepoints, adjusted for age, sex, mean FD (Model 1b)

@@

```{r, "qa check DMN model age,sex, mFD", echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

```{r "DMN2tp results age sex mFD", caption="DMN connectivity compared bl to fu, adjusted for age, sex and mFD", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.model.names = c("R0", "R1"), float.pos = "h",  custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Group BARS", "Age", "Sex", "logmFD", "Timepoint x Group"),  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

```{r "calc DMN three tp", eval = TRUE, echo = FALSE}
#### Aggregated DMN for three timepoints, adjusted for age, sex, mean FD (Model 1b)
R1 = lme4::lmer(mean_DMN_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_DMN_conn ~ tp + group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
anvDMN3tp=anova(R1,R0)
```

```{r "DMN3tp results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

```{r "calc rew two tp, age, sex", eval = TRUE, echo = FALSE}
### Reward network for two time points, adjusted for age and sex (Model 1a)

@@

```{r, "qa check Rew model age,sex", echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

```{r "Rew2tp results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex", "Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

```{r "calc rew two tp adjusted", eval = TRUE, echo = FALSE}
#### Reward network for two timepoints, adjusted for age, sex, mean FD (Model 1b)

@@

```{r "qa check Rew model age, sex, mFD", echo=FALSE}
diagnostics.plot(R0)
ranef.diagn.plot(R0)

@@

```{r "Rew2tpadj results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 7, 4, 5, 6),custom.coef.names=c("Intercept (BL, NBARS)","Timepoint FU", "Group BARS", "Age", "Sex","logmFD","Timepoint x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

```{r "calc rew three tp", eval = TRUE, echo = FALSE}
#### Reward network for three timepoints, adjusted for age, sex, mean FD (Model 1b)

R1 = lme4::lmer(mean_Rew_conn ~ tp*group_factor + Age_BL + Sex + logmFD +(1|subj.ID), data=final_FC)
R0 = lme4::lmer(mean_Rew_conn ~ tp + group_factor + Age_BL + Sex +logmFD +(1|subj.ID), data=final_FC)
anvRew3tp=anova(R1,R0)
```

```{r "Rew3tp results", caption="DMN connectivity compared bl to fu", results="asis", eval=FALSE}
#tab_model(R1,R0)
texreg(list(R0,R1), reorder.coef=c(1, 2, 3, 4, 8,9,5,6,7),custom.coef.names=c("Intercept (BL, NBARS)", "Timepoint FU", "Timepoint FU2", "Group BARS", "Age", "Sex","logmFD","Timepoint FU x Group", "Timepoint FU2 x Group"), custom.model.names = c("R0", "R1"), float.pos = "h",  center = TRUE, digits = 3, leading.zero = FALSE, single.row = TRUE, include.adjrs = FALSE, include.bic = FALSE, table = FALSE, use.packages = FALSE)
```

@@

#### Aggregated FC analysis
```{r "DMN model, mean, change BMI", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
p_meanBMI=anvmeanBMI$`Pr(>Chisq)`[2]

R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + Age_BL + Sex + logmFD + (1|subj.ID), data=final_FC)
R02a = lme4::lmer(mean_DMN_conn ~ within.BMI +  Age_BL + Sex + logmFD+(1|subj.ID), data=final_FC)
anvmeanBMI_adjmFD=anova(R1,R02a)
```
Corresponding to the whole brain results, there was a negative association of mean BMI and aggregated DMN connectivity (p = `r round(p_meanBMI,3)`), regardless of whether we adjusted for mFD.
```{r "Rew model, mean, change BMI", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_Rew_conn ~ within.BMI +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
```

```{r "DMN model mean+change of BMI+logmFD", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_DMN_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_DMN_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_DMN_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_DMN_conn ~ mean.BMI  + within.BMI +  mean.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
anvmeanFD=anova(R1,R03)
anvwithinFD=anova(R1,R03)

@@

#### Reward network, mean and within variability in BMI and logmFD
```{r "Rew model mean&change of BMI&logmFD", eval = TRUE, echo = FALSE}
R1 = lme4::lmer(mean_Rew_conn ~  mean.BMI + within.BMI + mean.logmFD + within.logmFD + Age_BL + Sex  + (1|subj.ID), data=final_FC)
R01 = lme4::lmer(mean_Rew_conn ~ mean.BMI + mean.logmFD + within.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
R02 = lme4::lmer(mean_Rew_conn ~ within.BMI + mean.logmFD + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
R03 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + mean.logmFD + Age_BL + Sex +(1|subj.ID), data=final_FC)
R04 = lme4::lmer(mean_Rew_conn ~ mean.BMI + within.BMI + within.logmFD +  Age_BL + Sex +(1|subj.ID), data=final_FC)
anvwithinBMI=anova(R1,R01)
anvmeanBMI=anova(R1,R02)
anvwithinFD=anova(R1,R03)
anvmeanFD=anova(R1,R04)
